Modeling and analyzing the Corona-virus warning app with the Isabelle Infrastructure framework

### 2020-07-10 paper.tex
a few more lines to complete the Refinement section

\subsection{Property Preserving System Refinement}
A first direct consequence of this definition is the following lemma
where the operator \texttt{$\ttimg$} in \texttt{\ttecal\ttimg(init K')}
represents function image, that is the set, $\{\ttecal(x). x \in \texttt{init K'}\} $.
\begin{ttbox}
{\bf{lemma}} init_ref: K \ttmeref K' \ttImp \ttecal\ttimg(init K') \ttsubseteq init K
\end{ttbox}
A more prominent consequence of the definition of refinement 
is that of property preservation. Here, we show that refinement preserves the
CTL property of ${\sf EF} s$ which means that a reachability property true in the
refined  model \texttt{K'} %has also been 
is already true in the abstract model.
A state set $s'$ represents a property %since we use 
in the predicate transformer view of properties as sets of states. 
The additional condition on initial states ensures that we cannot ``forget'' them. 
%initial states.
%The theorem states that, if the 
%property \texttt{{\sf EF} s'} is true in the concrete model its  has been true 
%can be reached from the initial state in 
%\texttt{K'} then it is also possible  to reach the image of this property
% in the abstract Kripke structure \texttt{K}.
\begin{ttbox}
{\bf{theorem}} prop_pres: 
   K \ttmeref K'  \ttImp init K \ttsubseteq \ttecal\ttimg(init K') \ttImp
   \ttforall s' \ttin Pow(states K'). K' \ttvdash {\sf EF} s' \ttimp K \ttvdash {\sf EF} (\ttecal\ttimg(s'))
\end{ttbox}
It is remarkable, that our definition of refinement by Kripke 
structure refinement entails property preservation and makes it possible 
to prove this as a theorem in Isabelle once for all, i.e., as a meta-theorem.
However, this is due to the fact that our generic definition of state transition
allows to explicitly formalize such sophisticated concepts like reachability.
For practical purposes, however, the proof obligation of showing that
a specific refinement is in fact a refinement is rather complex
justly because of the explicit use of the transitive closure of the state
transition relation.
In most cases, the refinement will be simpler. Therefore, we offer
additional help by the following theorem that uses a stronger characterization
of Kripke structure refinement and shows that our refinement follows
from this.
\begin{ttbox}
{\bf{theorem}} strong_mt: 
\ttecal\ttimg(init K') \ttsubseteq init K \ttand s \ttrel{\sigma'} s' \ttimp \ttecal(s) \ttrel{\sigma} \ttecal(s') 
\ttImp K \ttmeref K'
\end{ttbox}
This simpler characterization is in fact a stronger one: we could have $s \ttrel{\sigma'} s'$ 
in the refined Kripke structure \texttt{K'} and $\neg(\ttecal(s) \ttrel{\sigma} \ttecal(s'))$
but neither $s$ nor $s'$ are reachable from initial states in \texttt{K'}.
For cases, where we want to have the simpler one-step proviso but still need 
reachability we provide a slightly weaker version of \texttt{strong\_mt}.
\begin{ttbox}
{\bf{theorem}} strong_mt':  
\ttecal\ttimg(init K') \ttsubseteq init K \ttand (\ttexists s0 \ttin init K'. s0  \ttrelIstar s)
 \ttand s \ttrel{\sigma'} s' \ttimp \ttecal(s) \ttrel{\sigma} \ttecal(s') \ttImp K \ttmeref K'
\end{ttbox}

This idea of property preservation coincides with the classical idea of
trace refinement as it is given in process algebras like CSP. In this view,
the properties of a system are given by the set of its traces. Now, a refinement
of the system is given by another system that has a subset of the traces of the 
former one.
Although the principal idea is similar, we greatly extend it since our notion
additionally incorporates refinement. Since we include a state map 
\texttt{\ttsigma'\ttfun \ttsigma} in our refinement map, we additionally
allow structural refinement: the state map generalizes the basic idea of
trace refinement by traces corresponding to each other but allows additionally
an exchange of data types. 
As we see in the application to the case study, the refinement steps may
sometimes just specialize the traces: in this case the state map 
\texttt{\ttsigma'\ttfun \ttsigma} is just identity.

In addition, as we can observe in the following example in the second refinement
step, we also have a simple implicit version of {\it action refinement}. In an
action refinement, traces may be refined by combining consecutive system events
into atomic events thereby reducing traces. This will be detailed in the second refinement step
of the Corona-virus warning app considered next.

@@

\subsubsection{Second refinement step}
The persistent attack can be abbreviated informally by the action sequence \texttt{[get,move,move,get]}
performed by actors Eve, Bob, Eve, and Eve again, respectively.
How can a second refinement step avoid that Eve does the last \texttt{get} by imposing that after the
first \texttt{move} of Bob a \texttt{put} action must happen before Eve can do another \texttt{get}?
A very simple remedy to exclude this attack is to bind a \texttt{put} action after every \texttt{move}.
We can implement that change by a minimal update to the function \texttt{move\_graph\_a} (see Section \ref{sec:model})
by adding an increment (highlighted in the code snippet)
of the currently used Ephemeral Id before updating the \texttt{egre} component of the target location.
\begin{ttbox}
move\_graph\_a n l l' g \ttequiv ...
  (l' := insert (efids_cur({\bf efids\_inc\_ind}(efemid (cgra g n))))(egra g l))...
\end{ttbox}

>>>>

\subsection{Property Preserving System Refinement}
A first direct consequence of this definition is the following lemma
where the operator \texttt{$\ttimg$} in \texttt{\ttecal\ttimg(init K')}
represents function image, that is the set, $\{\ttecal(x). x \in \texttt{init K'}\} $.
\begin{ttbox}
{\bf{lemma}} init_ref: K \ttmeref K' \ttImp \ttecal\ttimg(init K') \ttsubseteq init K
\end{ttbox}
A more prominent consequence of the definition of refinement 
is that of property preservation. Here, we show that refinement preserves the
CTL property of ${\sf EF} s$ which means that a reachability property true in the
refined  model \texttt{K'} %has also been 
is already true in the abstract model.
A state set $s'$ represents a property %since we use 
in the predicate transformer view of properties as sets of states. 
The additional condition on initial states ensures that we cannot ``forget'' them. 
%initial states.
%The theorem states that, if the 
%property \texttt{{\sf EF} s'} is true in the concrete model its  has been true 
%can be reached from the initial state in 
%\texttt{K'} then it is also possible  to reach the image of this property
% in the abstract Kripke structure \texttt{K}.
\begin{ttbox}
{\bf{theorem}} prop_pres: 
   K \ttmeref K'  \ttImp init K \ttsubseteq \ttecal\ttimg(init K') \ttImp
   \ttforall s' \ttin Pow(states K'). K' \ttvdash {\sf EF} s' \ttimp K \ttvdash {\sf EF} (\ttecal\ttimg(s'))
\end{ttbox}
It is remarkable, that our definition of refinement by Kripke 
structure refinement entails property preservation and makes it possible 
to prove this as a theorem in Isabelle once for all, i.e., as a meta-theorem.
However, this is due to the fact that our generic definition of state transition
allows to explicitly formalize such sophisticated concepts like reachability.
For practical purposes, however, the proof obligation of showing that
a specific refinement is in fact a refinement is rather complex
justly because of the explicit use of the transitive closure of the state
transition relation.
In most cases, the refinement will be simpler. Therefore, we offer
additional help by the following theorem that uses a stronger characterization
of Kripke structure refinement and shows that our refinement follows
from this.
\begin{ttbox}
{\bf{theorem}} strong_mt: 
\ttecal\ttimg(init K') \ttsubseteq init K \ttand s \ttrel{\sigma'} s' \ttimp \ttecal(s) \ttrel{\sigma} \ttecal(s') 
\ttImp K \ttmeref K'
\end{ttbox}
This simpler characterization is in fact a stronger one: we could have $s \ttrel{\sigma'} s'$ 
in the refined Kripke structure \texttt{K'} and $\neg(\ttecal(s) \ttrel{\sigma} \ttecal(s'))$
but neither $s$ nor $s'$ are reachable from initial states in \texttt{K'}.
For cases, where we want to have the simpler one-step proviso but still need 
reachability we provide a slightly weaker version of \texttt{strong\_mt}.
\begin{ttbox}
{\bf{theorem}} strong_mt':  
\ttecal\ttimg(init K') \ttsubseteq init K \ttand (\ttexists s0 \ttin init K'. s0  \ttrelIstar s)
 \ttand s \ttrel{\sigma'} s' \ttimp \ttecal(s) \ttrel{\sigma} \ttecal(s') \ttImp K \ttmeref K'
\end{ttbox}
This idea of property preservation coincides with the classical idea of
trace refinement as it is given in process algebras like CSP. In this view,
the properties of a system are given by the set of its traces. Now, a refinement
of the system is given by another system that has a subset of the traces of the 
former one.
Although the principal idea is similar, we greatly extend it since our notion
additionally incorporates refinement. Since we include a state map 
\texttt{\ttsigma'\ttfun \ttsigma} in our refinement map, we additionally
allow structural refinement: the state map generalizes the basic idea of
trace refinement by traces corresponding to each other but allows additionally
an exchange of data types. 
As we see in the application to the case study, the refinement steps may
sometimes just specialize the traces: in this case the state map 
\texttt{\ttsigma'\ttfun \ttsigma} is just identity.

In addition, we also have a simple implicit version of {\it action refinement}. In an
action refinement, traces may be refined by combining consecutive system events
into atomic events thereby reducing traces.
We can observe this kind of refinement in the second refinement step
%This will be detailed in the second refinement step
of the Corona-virus warning app considered next.

@@

\subsubsection{Second refinement step}
The persistent attack can be abbreviated informally by the action sequence \texttt{[get,move,move,get]}
performed by actors Eve, Bob, Eve, and Eve again, respectively.
How can a second refinement step avoid that Eve does the last \texttt{get} by imposing that after the
first \texttt{move} of Bob a \texttt{put} action must happen before Eve can do another \texttt{get}?
A very simple remedy to exclude this attack is to bind a \texttt{put} action after every \texttt{move}.
We can implement that change by a minimal update to the function \texttt{move\_graph\_a} (see Section \ref{sec:model})
by adding an increment (highlighted in the code snippet)
of the currently used Ephemeral Id before updating the \texttt{egre} component of the target location.
\begin{ttbox}
move\_graph\_a n l l' g \ttequiv ...
  (l' := insert (efids_cur({\bf efids\_inc\_ind}(efemid (cgra g n))))(egra g l))...
\end{ttbox}  
This is an action refinement because the move action is changed. It is a refinement, since
any trace of the refined model can still be mapped to a trace in the more abstract model just
omitting a few steps (the refinement relation is defined using the reflexive transitive closure
\texttt{\ttrelIstar}.
